---
title: "HW 12: Dates"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Load the tidyverse, here and lubridate libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
```

## Part 1: Play Date

```{r}
birth_date <- "11/25/1998"

# Elizabeth Garret Anderson was the first female physician
anderson <- "9 June 1836"

# Mary Anning discovered the first complete Plesiosaur fossil, 
# and other important Jurassic fossils
anning <-  "1799/05/21"

# Alice Augusta Ball was a chemist who developed the
# first injectable treatment for leprosy.
ball <- "July 24, 1892"

# Sylvia Earle was the first chieft scientist at NOAA
earle <- "August, 30 (1935)"

# Rosalind Franklin used x-ray diffraction to 
# discover DNA structure. Watson and Crick claimed credit.
franklin <- "25th of July, 1920"

# Ester Lederberg was the first to discover that viruses 
# infect bacteria. Led to discovery of lambda bacteriophage
lederberg <- "December 18, 1922"

# Barbara McClintock discovered transposable elements (jumping genes),
# and developed staining techniques to study chromosomes
mcclintock <- "16061902"

# Let's also remember a dream
mlk_birth <- "1/15/1929"
mlk_dream <- "1963 August 28"
```

#### Answer the following questions using code chunks

1. Convert each date to POSIXct format, using lubridate functions, saving each to a unique variable. Print each result.

```{r}
birth_date <- mdy(birth_date)
birth_date

anderson_birth_date <- dmy(anderson)
anderson_birth_date

anning_birth_date <- ymd(anning)
anning_birth_date

ball_birth_date <- mdy(ball)
ball_birth_date

earle_birth_date <- mdy(earle)
earle_birth_date

franklin_birth_date <- dmy(franklin)
franklin_birth_date

lederberg_birth_date <- mdy(lederberg)
lederberg_birth_date

mcclintock_birth_date <- dmy (mcclintock)
mcclintock_birth_date

mlk_birth_date <- mdy(mlk_birth)
mlk_birth_date

mlk_dream_date <-ymd(mlk_dream)
mlk_dream_date
```
2. Check yours dates to be sure they are correct. If you spot any errors, correct them now using the proper lubridate function(s).
-All dates check out

3. What day of the week were you born on? Show the full day name, not the abbreviated name or the day number.

```{r}
wday(birth_date, label = TRUE, abbr = FALSE)
```
I was born on a Wednesday. 

4. What day of the week will your birthday be on this year? Use lubridate’s update function. You must figure out how to do it entirely by code for full points (no hard-coding). In other words, you code should work just as well next year or five years in the future, without editing.

```{r}
update(birth_date, year = 2020) %>% 
  weekdays(abbreviate = FALSE)
```
My birthday will be on a Wednesday again. 


5. What day of the week would your birthday be on in the Year 2525, if Man is still alive? How old would you be? Here you may hard-code 2525. Show the abbreviated day of the week.

```{r}
age <- dmy(25112525) - birth_date
as.duration(age)
update(birth_date, year = 2525) %>%
  weekdays(abbreviate = TRUE)
```
In the year 2525, I would be 527 years old, and my birthday would be on a Sunday. Given that humans still inhabit the earth. 


6. How many days was it between the birth of Dr. Elizabeth Garrett Anderson and Alice Augusta Ball?

```{r}
ball <- ball_birth_date - anderson_birth_date
print(ball)
```
There are 20,499 days between their birth.


7. What is the duration between Martin Luther King Jr.’s birth date and his I Had a Dream speech.

```{r}
mlk_dream_date - mlk_birth_date
```
There is a difference of 57,526 days.


8. Enter the eight POSIXct compatible birth dates, including your own, into a vector in the order they are given above. Name each element of the vector with the last name of each person. 

```{r}
dates <- c(birth_date, anderson_birth_date, anning_birth_date, ball_birth_date, earle_birth_date, lederberg_birth_date, mcclintock_birth_date, franklin_birth_date, mlk_birth_date)
names(dates) <- c('Laubinger','Anderson', 'Anning', 'Ball', 'Earle', 'Lederberg', 'McClintock', 'Franklin', 'King')
leap_year(dates)
difference <- dates - birth_date
print (print(difference))
```


I think this is correct. I am listed last, with 0 days. 


## Part 2: Wrangling and plotting with dates

### import lake_ice.csv 
```{r}
lake_ice_data <- read_csv(here::here("data", "lake_ice.csv"),na="-999") %>%
 na.omit()
lake_ice_data
```

### Filter data to save only Lake Mendota
```{r}
lake_mendota_data <- filter(lake_ice_data, lakename == "LAKE MENDOTA")
```


### Select these six columns. You can do this in one of a few different ways but you should be able to do this very efficiently
```{r}
lake_mendota_data_tidy <- select(lake_mendota_data, iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day) %>%
  mutate(ice_on = make_date(iceon_year, iceon_month, iceon_day) ,ice_off = make_date(iceoff_year, iceoff_month, iceoff_day), duration_days = as.numeric(ice_off-ice_on))
```

```{r}
lake_mendota_data_tidy
```


### Make a line graph to show how duration changed over iceoff_year

```{r}
ggplot(data = lake_mendota_data_tidy)+
  geom_line(mapping = aes(x=iceoff_year, y=duration_days))+xlab("Year")+ylab("Duration of Ice(Days)")
```
I think this is correct. It matches the graph in the assignment. 





















