---
title: "HW05 Part 2: Data Visualization"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

##LOad Tidyverse
```{r}
library(tidyverse)
```
#### Plot 1

Load dataset

```{r}
data(trees)
```

```{r}
?trees
```

Create the scatterplot

```{r}
ggplot(data = trees)+
  geom_point(mapping = aes(x=Height, y=Girth))
```

As height increases, so does girth. As height decreases, so does girth. 

#### Plot 2

Load "lynx" dataset

```{r}
data("lynx")
```

```{r}
?lynx
```

Make a vector called year
```{r}
year <- c(1821:1934)
```

Look at class() of lynx dataset
```{r}
class(lynx)
```

Convert the time series data to a vector

```{r}
pelts <- as.vector(lynx)
```


Make a dataframe called "lynx_pelts"

```{r}
lynx_pelts <- data.frame(year, pelts)
```

Make a line graph plot

```{r}
ggplot(data = lynx_pelts)+
  geom_smooth(mapping = aes(x = year, y = pelts, color= "maroon"))
```

This shows the years lynx pelts were collected and the number they were collect. Pelt collection has been on the decline since around 1920. 


#### Plot 3

Load "iris" dataset

```{r}
data("iris")
```

```{r}
?iris
```

Make plot with x = petal length and y = petal width

```{r}
ggplot(data = iris, mapping = aes(x = Petal.Length, y = Petal.Width)) + labs(y = "Petal Width", x = "Petal Length") + geom_point(mapping = aes(color = Species, shape = Species, point = 2))
```

This shows the three species of iris, color coated. Setosa have the smallest petal width and length, while virginica has the largest. 



#### plot 4

Load " msleep" dataset

```{r}
data("msleep")
```


```{r}
?msleep
```

Make a violin plot of "msleep data"

```{r}
ggplot(data = msleep, mapping = aes(x = vore, y = sleep_rem))+
  geom_violin(fill = "gray30")
```


Few carnivores are able to have the highest amount of rem sleep, but the data is varied. Most animals have at least 2 hours of rem sleep. 

#### Plot 5

Make another Violin plot

```{r}
ggplot(data = msleep) +
  geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "gray70")
```

This shows that same data as plot 4, just in a lighter shade of gray. Violin plots allow for all data to be arranged, and shows the median ranges. 

#### plot 7

Using same data in plot 4 & 5 to make a boxplot

```{r}
ggplot(data = msleep) +
  geom_boxplot(mapping = aes(x = order, y = sleep_total)) + 
  coord_flip()
```

This data is more scattered and difficult to read with the large number of orders. There is less of a pattern to distinguish, than that of plot 4 & 5.


#### Plot 8 

Using "msleep" data to make a boxplot with points

```{r}
ggplot(data = msleep, mapping = aes(x = conservation, y = awake, color = conservation)) +
  geom_boxplot() +
  geom_point() +
  geom_jitter() +
  coord_flip() +
  labs(x = 'Conservation', y = 'Awake') +
  scale_color_discrete('Conservation')
  
```
This data shows the boxplot with points, and is easier to see than the traditional boxplot in plot 7. Most are awake from 5 to 10 hours. 


#### plot 9 

Make a scatterplot of awake variables against body weight.

```{r}
ggplot(data = msleep) +
  geom_point(mapping = aes(x = awake, y = bodywt, color = conservation)) +
  facet_wrap(~ vore, nrow = 2)
```

The values for bodywt make it difficult to see data below 2000. Herbivores are the only ones with a body weight above 2000 and are awake over 20 hours. 

#### Plot 10

Make a scatterplot with awake 
variables against brain weight.

```{r}
ggplot(data = msleep) +
  geom_point(mapping = aes(x = awake, y = brainwt, color = vore))
```

This shows that herbivores have the largest brain weight, and thus are awake longer than those that have smaller brains. 