---
title: "HW05 Part 2: Data Visualization"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

##LOad Tidyverse
```{r}
library(tidyverse)
```
#### Plot 1

##Load dataset

```{r}
data(trees)
```

```{r}
?trees
```


## Create the scatterplot

```{r}
ggplot(data = trees)+
  geom_point(mapping = aes(x=Height, y=Girth))
```

As height increases, so does girth. As height decreases, so does girth. 

#### Plot 2

## Load "lynx" dataset

```{r}
data("lynx")
```

```{r}
?lynx
```

## Make a vector called year
```{r}
year <- c(1821:1934)
```

## Look at class() of lynx dataset
```{r}
class(lynx)
```

## Convert the time series data to a vector

```{r}
pelts <- as.vector(lynx)
```


## Make a dataframe called "lynx_pelts"

```{r}
lynx_pelts <- data.frame(year, pelts)
```

```{r}
ggplot(data = lynx_pelts)+
  geom_smooth(mapping = aes(x = year, y = pelts, color= "maroon"))
```

This shows the years lynx pelts were collected and the number they were collect. Pelt collection has been on the decline since around 1920. 


#### Plot 3

## Load "iris" dataset

```{r}
data("iris")
```

```{r}
?iris
```

## Make plot with x = petal length and y = petal width

```{r}
ggplot(data = iris, mapping = aes(x = Petal.Length, y = Petal.Width)) + labs(y = "Petal Width", x = "Petal Length") + geom_point(mapping = aes(color = Species, shape = Species, point = 2))
```

This shows the three species of iris, color coated. Setosa have the smallest petal width and length, while virginica has the largest. 

















