---
title: "HW09: Wrangling II"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load all the packages 

```{r}
library(tidyverse)
```

```{r}
library(here)
```

```{r}
library(smatr)
```

```{r}
library(knitr)
```


## Use Rsource to load functions
```{r}
my_functions <- source(here('scripts', 'my_functions.R'))
```

### Darters in riffles

## import data
Read in darters.csv. These are the same data as an earlier assignment but in csv format.
Use filter() to remove tetrazonum and zonale.
Do not delete the minsub column like you did in the previous assignment.

```{r}
darters_data <- read_csv(here("data", "darters.csv"))
darters_data <- filter(darters_data, species != "tetrazonum" & species != "zonale")
darters_data
```
This code chunk imports the data from darters.csv found in the data folder


## Wrangle Data
Use mutate() to perform these wrangling tasks 
Only total_substrate needs a new column.

```{r}
wrangled_darters <- darters_data %>% 
  mutate(sex = ifelse(sex == "f",                              
                             "female", 
                             "male"),
         total_substrate = majsub + minsub,
         length = length * 10,
         riffle = ifelse(riffle == "1",
                         "riffle 1",
                         "riffle 2"))
wrangled_darters
```
This code chunk has summed  together the majsub and minsub into a new column called total_substrate, changed riffle values 1 and 2 into Riffle 1 and Riffle 2, changed length from centimeter to millimeter, and sex from f to female and m to male.

## Does total_substrate sum to 1 for all observations?

```{r}
wrangled_darters$total_substrate == 1
```
All are true for equaling 1. 


## Summarize Data
Use group_by() and summarize() to summarize the mean length, depth, and velocity for each species for each riffle and use facet_wrap on riffle

```{r}
(darters_summary <- wrangled_darters %>%
  group_by(riffle, species) %>%
  summarise(length = mean(length, na.rm = TRUE),
            depth = mean(depth, na.rm = TRUE),
            velocity = mean(velocity, na.rm = TRUE)))
```

## Use kable to make a table with summary means
```{r}
kable(darters_summary)
```
 This gives a summary of the length, depth, and velocity of species. 
 
## Graph data

## Plot 1 
Make a plot that uses the stat_summary() technique and use x = species and y = length.

```{r}
wrangled_darters %>%
   filter(!is.na(length)) %>%
  ggplot(mapping = aes(x = species, y = length)) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
    ) +
   geom_point(data = darters_summary,
             mapping = aes(x = species,
                           y = length),
             shape = 15,
             color = "yellow") +
  facet_wrap(~ riffle) +
  labs(y = "Length(in mm)", x = NULL)+
  coord_flip()
```
Spectabile seems to have the largest range, from around 320 mm to 355mm, while biennioides seems to have the smallest, with around 350mm to 360mm.Specie biennioides has the largest size in fish in riffle 2, near 360mm.In spectabile and caeruleum, they have the larger sized fish in riffle 1. 

## Plot 2
Make a new plot with a geom_point() layer of the length variable for each species. Change the point color to a lighter shade in the range of gray50 to gray70. Add a stat_summary() layer. Add another stat_summary(). Change the y function to mean, size to 3, and add the color = "maroon" argument.Facet, label, and flip the graph as you did above.

```{r}
wrangled_darters %>%
   filter(!is.na(length)) %>%
  ggplot(mapping = aes(x = species, y = length)) +
  geom_point(color = "gray55") +
  stat_summary(fun.y = median,
               geom = "point",
               size = 2.2) +
  stat_summary(fun.y = mean,
               geom = "point",
                size = 3,
                color = "maroon") +
  facet_wrap(~ riffle) +
  labs(y = "Length(in mm)", x = NULL)+
  coord_flip()
```
Changing the mean to size 3 and maroon, allows for it to be more easily seen , compared to the other data points that are grey55 and size 2.2. This lets the data for riffle 1 and 2 to be more easily seen for the mean over median. 



### Stress and corticosterones in birds

## Load libraries 
```{r}
library(tidyverse)
```
```{r}
library(smatr)
```

## import data file

```{r}
feathers_data <- read_csv(here("data", "feathers.csv"))
```

## use 'locale' argument to specifiy decimal mark 

```{r}
?locale
```
```{r}
locale(feathers_data = "en", read.csv,
  decimal_mark = ".", grouping_mark = ",", tz = "UTC",
  encoding = "UTF-8", asciify = FALSE)
```




























