---
title: "HW09: Wrangling II"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load all the packages 

```{r}
library(tidyverse)
```

```{r}
library(here)
```

```{r}
library(smatr)
```

```{r}
library(knitr)
```


## Use Rsource to load functions
```{r}
my_functions <-(here('scripts', 'my_functions.R'))
```

### Darters in riffles

## import data
Read in darters.csv. These are the same data as an earlier assignment but in csv format.
Use filter() to remove tetrazonum and zonale.
Do not delete the minsub column like you did in the previous assignment.

```{r}
darters_data <- read_csv(here("data", "darters.csv"))
darters_data <- filter(darters_data, species != "tetrazonum" & species != "zonale")
darters_data
```
This code chunk imports the data from darters.csv found in the data folder


## Wrangle Data
Use mutate() to perform these wrangling tasks 
Only total_substrate needs a new column.

```{r}
wrangled_darters <- darters_data %>% 
  mutate(sex = ifelse(sex == "f",                              
                             "female", 
                             "male"),
         total_substrate = majsub + minsub,
         length = length * 10,
         riffle = ifelse(riffle == "1",
                         "riffle 1",
                         "riffle 2"))
wrangled_darters
```
This code chunk has summed  together the majsub and minsub into a new column called total_substrate, changed riffle values 1 and 2 into Riffle 1 and Riffle 2, changed length from centimeter to millimeter, and sex from f to female and m to male.

## Does total_substrate sum to 1 for all observations?

```{r}
wrangled_darters$total_substrate == 1
```
All are true for equaling 1. 


## Summarize Data
Use group_by() and summarize() to summarize the mean length, depth, and velocity for each species for each riffle and use facet_wrap on riffle

```{r}
(darters_summary <- wrangled_darters %>%
  group_by(riffle, species) %>%
  summarise(length = mean(length, na.rm = TRUE),
            depth = mean(depth, na.rm = TRUE),
            velocity = mean(velocity, na.rm = TRUE)))
```

## Use kable to make a table with summary means
```{r}
kable(darters_summary)
```
 This gives a summary of the length, depth, and velocity of species. 
 
## Graph data

## Plot 1 
Make a plot that uses the stat_summary() technique and use x = species and y = length.

```{r}
wrangled_darters %>%
   filter(!is.na(length)) %>%
  ggplot(mapping = aes(x = species, y = length)) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
    ) +
   geom_point(data = darters_summary,
             mapping = aes(x = species,
                           y = length),
             shape = 15,
             color = "yellow") +
  facet_wrap(~ riffle) +
  labs(y = "Length(in mm)", x = NULL)+
  coord_flip()
```
Spectabile seems to have the largest range, from around 320 mm to 355mm, while biennioides seems to have the smallest, with around 350mm to 360mm.Specie biennioides has the largest size in fish in riffle 2, near 360mm.In spectabile and caeruleum, they have the larger sized fish in riffle 1. 

## Plot 2
Make a new plot with a geom_point() layer of the length variable for each species. Change the point color to a lighter shade in the range of gray50 to gray70. Add a stat_summary() layer. Add another stat_summary(). Change the y function to mean, size to 3, and add the color = "maroon" argument.Facet, label, and flip the graph as you did above.

```{r}
wrangled_darters %>%
   filter(!is.na(length)) %>%
  ggplot(mapping = aes(x = species, y = length)) +
  geom_point(color = "gray55") +
  stat_summary(fun.y = median,
               geom = "point",
               size = 2.2) +
  stat_summary(fun.y = mean,
               geom = "point",
                size = 3,
                color = "maroon") +
  facet_wrap(~ riffle) +
  labs(y = "Length(in mm)", x = NULL)+
  coord_flip()
```
Changing the mean to size 3 and maroon, allows for it to be more easily seen , compared to the other data points that are grey55 and size 2.2. This lets the data for riffle 1 and 2 to be more easily seen for the mean over median. 



### Stress and corticosterones in birds

## Load libraries 
```{r}
library(tidyverse)
```
```{r}
library(smatr)
```

## import data file

```{r}
?locale
```

```{r}
 feathers_data <- read_csv(here("data","feathers.csv"),locale = locale(decimal_mark = ",")) %>% 
  rename(capture_date="Capture Date",day=Date,site=Site,urban_score="Urba Score",cortF="CortF ng/mg",cortB="CortF ng/mm", sex=Sex,smi=SMI,tarsus=Tarsus,mass=Mass,basal_cort=Bas_Cort,stress_cort=Stress_Cort)
feathers_data
```
The code chunk imported the data, changed the decimal_marks to "," , searched the locale help, and changed column names including spaces and units of measurement. 

## Wrangle Data
#Calculate the mean tarsus length of all individuals in the data set, and store it in a variable called mean_tarsus.

```{r}
mean_tarsus <- mean(tarsus,na.rm = TRUE)
mean_tarsus
```

#Use mutate() to log-transform (log()) the body mass (mass) and the tarsus length (tarsus)

```{r}
mut_feathers <- mutate(feathers_data, log_mass=log(mass),log_tarsus=log(tarsus))
mut_feathers
```

# Run major_axis <- sma(log_mass ~ log_tarsus, data = <data>). Substitute the name of your imported data for <data>
```{r}
major_axis <- sma (log_mass ~ log_tarsus, data=mut_feathers)
```

# Apply the summary() function to major_axis

```{r}
major_axis <- sma (log_mass ~ log_tarsus, data=mut_feathers)
summary(major_axis)
```
# Apply the coef() function to major_axis. 
```{r}
coef(major_axis)
major_axis_coef <- coef(major_axis)
```

# Store value in varibale 'b'. 
```{r}
b <- major_axis_coef['slope']
b
```
A slope of 1.864 is saved for 'b'.

# Use mutate and scaled_mass function to calculate SMI in new column called new_smi
```{r}
smi_mut_feathers <- mutate(mut_feathers,new_smi=scaled_mass(mass,tarsus,b))  
smi_mut_feathers
```

# Compare your new column with the column you imported

```{r}
select(smi_mut_feathers,smi,new_smi)
```
They are nearly identical, off by around .01 on some. 

## Summarize data
#group by site and use summarize() to calculate mean and std_err of mean.
```{r}
sum_feathers <- group_by(smi_mut_feathers,site)
feathers_sum <- summarise(sum_feathers,
                          basal_mean =  mean(basal_cort,na.rm=TRUE),
    stress_mean = mean(stress_cort,na.rm=TRUE),
    std_err_basal=std_err(basal_cort,na.rm=TRUE),
    std_err_cort=std_err(stress_cort,na.rm=TRUE),
    .groups='drop')
feathers_sum
```
This has created columns for basal_mean, stress_mean, std_err_basal, and std_err_cort.

## Graph data
#Plot 1: make a box plot of new_smi for each site. 
```{r}
ggplot(data=smi_mut_feathers)+
  geom_boxplot(mapping=aes(x=site,y=new_smi))
```
All sites have similar medians, around 25, except villefollet which has a median new 23. As for range, NIORT has the largest range by far.

# Plot 2: A common graph is to plot the means with error bars based on the standard errors of the means.Make a plot of basal cortiocosterone concetration for each site. 

```{r}
ggplot(data=feathers_sum)+
  geom_point(mapping = aes(x=site,y=std_err_basal))+
  geom_errorbar(mapping=aes(x=site,ymin=std_err_basal-1,ymax=std_err_basal+1))
```
This shows the standard error of the basal concentration. CEBC and VILLEFOLLET are simialar, around 0.4, while LAROCHELLE is at 0.5 and NIORT is near 0.7.

# Plot 3: 
```{r}
ggplot(data=feathers_sum)+
  geom_point(mapping=aes(x=site,y=std_err_cort))+
  geom_linerange(mapping = aes(x=site,ymin=std_err_cort-1,ymax=std_err_cort+1))
```
This code chunk removes the horizontal default error lines, removing the tiefighter look of the previous plot. 


#### Vóila































