---
title: "HW 10: Data Visualization 1"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Install packages
```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```

### Import and Wrangle 

## import sparrow.txt.
```{r}
sparrows_raw <- read_tsv(here("data", "sparrows.txt"))
```
```{r}
sparrows_raw
```
Data is checked and imported

## Wrangle sparrow.txt 
```{r}
sparrows <- sparrows_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))
```

```{r}
view(sparrows)
```


### Figure 2: Outliers

## Run code to generate boxplot
```{r}
plot_box <- sparrows %>%
  ggplot(aes(x = box_group, y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL, y = "Wing length (mm)")
```

# Run code associated with Cleveland dot plot
```{r}
plot_cleveland <- sparrows %>% 
  ggplot(aes(x = wingcrd,
             y = row_order)) +
  geom_point() +
  labs(x = "Wing length (mm)",
       y = "Order of the data")
```

## Display plot_box and plot_cleveland 
```{r}
plot_box + plot_cleveland
```
## Use coord_flip to aligne axes
```{r}
plot_box + plot_cleveland + coord_flip()
```
This causes wing length to remain on the same axes for both of the graphs. 

### Figure 3: Outliers

## p_culmen code chunk
```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=culmen))+
  theme(axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank())
```

## p_nalospi code chunk 
```{r}
p_nalospi <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=nalospi))+
  theme(axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank())
```

## p_wt code chunk
```{r}
p_wt <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=wt))+
  theme(axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank())
```

## p_wingcrd code chunk
```{r}
p_wingcrd <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=wingcrd))+
  theme(axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank())
```

## p_tarsus code chunk 
```{r}
p_tarsus <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=tarsus))+
  theme(axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank())
```

## p_head code chunk
```{r}
p_head <- sparrows %>% 
  ggplot(aes(y=row_order))+
  geom_point(aes(x=head))+
  theme(axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank())
```

## display code chunks together 
```{r}
p_culmen+p_nalospi+p_wt+p_wingcrd+p_tarsus+p_head +plot_layout(ncol=3,nrow=2)
```
Tarsus, nalospi, and head have the outlier data. 


### Figure 5: distribution

## Histogram 
```{r}
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
```

## Faceted Histogram
```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)
```

## Assemble two plots
```{r}
h1 + h2
```
### Figure 10: Relationships

## scatterplot with ggscatmat
```{r}
sparrows %>% ggscatmat(columns = 1:7)
```
This makes graphs much faster and requires less work. 

### Figure 11: interactions 

## Use factor() function to convert months to factors as in part of mutate()
```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex != 0) %>% 
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>% 
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month)
```
This creates trend lines in the panels. Males tend to go to the left, while females tend to trend to the right. It also lets you view the months you wanted to use from the data set. 


### The End :)



































