---
title: "HW 14: Maps"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load libraries 
```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

## Part 1: outline maps 

### U.S maps

# Outline of the 48 contiguous states
```{r}
usa <- map_data(map = "state")
```

## Outline of just Missouri
```{r}
us_missouri <- map_data(map = "state",
                        region = "missouri")
```

# Outline of Alabama, Florida, Mississippi, and North Carolina
```{r}
us_other <- subset(x = usa, 
                   subset = region %in% c("alabama",
                                          "florida",
                                        "mississippi",
                                    "north carolina"))
```

# lay the first layer of maps 
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray")
```

# add missouri to the map
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black")
```
Missouri has been added. 

# add outline of all 48 states
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black")
```
Full map, with Missouri filled and other 4 outlined.

# Add the filled states first so that the black outlines of all
# states are overlaid on top.
```{r}
us_map <- ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black") + 
  theme_map() +
  coord_fixed(1.3)

us_map
```


# build Missouri map and highlight Scott County
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <- ggplot() + 
  geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") +
  theme_map() +
  coord_fixed(1.3)

mo_map
```

# put Missouri map and us map side by side 
```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5,1))
```
checks out, two maps side by side 

### outline maps: my turn
# highlight missouri on the us map
```{r}
us_map2 <- ggplot() + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black")
us_map2
```

# highlight two counties in gray and birth county in black
```{r}
missouri <- map_data("county", "missouri")
mo_franklin <- subset(missouri, subregion %in% c("franklin"))
mo_holt <- subset(missouri, subregion %in% c("holt"))
mo_lewis <- subset(missouri, subregion %in% c("lewis"))
mo_map2 <- ggplot() + 
  geom_polygon(data = missouri, aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_franklin, aes(x = long, y = lat, group = group), fill = "black") + 
  geom_polygon(data = mo_holt, aes(x = long, y = lat, group = group), fill = "gray55") + 
  geom_polygon(data = mo_lewis, aes(x = long, y = lat, group = group), fill = "gray55") + 
  coord_fixed(1.3) + 
  theme_map()

mo_map2
```

#combine two maps side by side
```{r}
us_map2 + mo_map2 +plot_layout(ncol = 2, widths = c(1.5,1))
```
Bam! two maps of my birth state and county in black, as well as two additional counties in gray. 


## Part 2: Choropleth maps 

# load library
```{r}
library(sf)
library(tidyverse)
```

# import skin data 
```{r}
skin_cancer_raw <- 
  read_delim(file = here::here("data","skin_cancer.txt"), 
             delim = " ", 
             skip = 7, 
             trim_ws = TRUE)
```

# wrangle data 
```{r}
skin_cancer <- skin_cancer_raw %>% 
  filter(State != "Wash,D.C.") %>% 
  mutate(State = case_when(
    State == "NorthCarolina" ~ "North Carolina",
    State == "SouthCarolina" ~ "South Carolina",
    State == "NorthDakota" ~ "North Dakota",
    State == "SouthDakota" ~ "South Dakota",
    State == "NewHampshire" ~ "New Hampshire",
    State == "NewJersey" ~ "New Jersey",
    State == "NewMexico" ~ "New Mexico",
    State == "MewYork" ~ "New York", # Data has MewYork typo 
    State == "RhodeIsland" ~ "Rhode Island",
    State == "WestVirginia" ~ "West Virginia",
    TRUE ~ State
  ))
```

# calculate average mortality for all 48 states. Then use mutate() to add a relative_mort column, substracting the Mort column from the mean_mort average
```{r}
mean_mort <- mean(skin_cancer$Mort)

skin_cancer_adj <- skin_cancer %>% 
  mutate(relative_mort = Mort - mean_mort)
```

# Used to trim Alaska, Hawaii, and territories from the shapefile.
```{r}
lower_48 <- c("Alabama","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
```

# filter and only keep lower_48 states
```{r}
states <- st_read(here::here("data/cb_2017_us_state_500k.shp"),
                  stringsAsFactors = TRUE)
```
```{r}
states_df <- states %>%
  dplyr::filter(NAME %in% lower_48)
```

# create column called STATEFP
```{r}
str(states_df$STATEFP)
```

# Capture the FIPS levels from states_df
# Use to factor the FIPS and state names in the next step.
```{r}
state_levels <- levels(states_df$STATEFP)
name_levels <- levels(states_df$NAME)
```

```{r}
skin_cancer_adj <- left_join(skin_cancer_adj, fips)

states_df <- left_join(states_df, skin_cancer_adj)
```

#change the STATEFP column to an unordered factor.
```{r}
library(tidyverse)
fips <- read_csv(here::here("data","state_fips.csv")) %>% 
  select(-abbr) %>% 
  rename(STATEFP = FIPS) %>% 
  mutate(STATEFP = factor(STATEFP, levels = state_levels))
```

```{r}
ggplot(states_df) +
  geom_sf(aes(fill = relative_mort)) +
  scale_fill_viridis_c(name = "Skin Cancer,\nRelative Mortality",
                       option = "inferno") +
  coord_sf(crs = st_crs(5070)) +
  theme_bw() +
  theme(legend.position = "bottom")
```
It seems like skin cancer mortality is higher in the southern states, which is expected as they receive more intense sunlight closer to the equator .

## Choropleth map: your turn

# import data
```{r}
lyme_disease_raw<- read_csv(here::here("data", "lyme_disease.csv"),
                            skip = 1,
                            na = c("-")) 
lyme_disease_raw
```
use na = c("-") to remove missing data

# wrangle data, create new column and remove columns,  pivot_longer to put the years into a single column, group_by state and use summarize(),
```{r}
lyme_disease_tidy <- lyme_disease_raw %>%
  mutate(`2017` = `2017_Confirmed` + `2017_Probable`) %>% 
   select(State: "2016", "2017") %>% 
  pivot_longer("2007":"2017", names_to = "year", values_to = "cases")%>% 
  group_by(State) %>% 
  summarise(mean_cases = mean(cases), 
            log_mean = log10(mean_cases), 
            .groups = "drop")
```

# Use left_join as you did above to merge the fips data frame
```{r}
lyme_disease_adj <- left_join(lyme_disease_tidy, fips)
```
# Use left_join again to merge your lyme data to the states_df data
```{r}
lyme_disease_df <- left_join(states_df, lyme_disease_tidy)
```

# Plot your choropleth map using the same ggplot code for the cancer map, change aes(file = ...) for geom_sf, scale approprately, add labels 
```{r}
ggplot(lyme_disease_df)+
   geom_sf(aes(fill = log_mean)) +
  scale_fill_viridis_c(name = "Lyme Disease,\nMean Cases",
                       option = "viridis",
                       labels = c("<1", "10", "100", "1000", "5000")) +
  coord_sf(crs = st_crs(5070)) +
  theme_bw() +
  theme(legend.position = "bottom")
```
Map checks out with the notes. Yay!

## Part 3: Dot distribution maps

# load library
```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(ggmap)
```

# Global vars -------------------------------------------------------------

# Restrict the map to the western US.
```{r}
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(9618973)
```

# Bigfoot data ------------------------------------------------------------
```{r}
bigfoot <- read_csv(here::here("data", "bigfoot.csv"))

bigfoot <- bigfoot %>% 
  filter(long >= min_long & long <= max_long,
         lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  dplyr::select(name, long, lat)
```

# Bear data --------------------------------------------------------------
```{r}

bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```

```{r}
tail(both_species)
```


# get a terrain map of data 
```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat),
                     zoom = 4, 
                     maptype = "terrain-background")



ggmap(base)
```


# add plot point layer for big foot 
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```


# add point layer for black bear
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))
```

# make a final plot for both species 
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```


## make my own distribution map

# change setseed to last 7 digits of SO #
```{r}
min_long <- -126
max_long <- -104
min_lat <-  31
max_lat <- 50

set.seed(1977039)
```


# Bigfoot data 
```{r}
bigfoot <- read_csv(here::here("data", "bigfoot.csv"))

bigfoot <- bigfoot %>% 
  filter(long >= min_long & long <= max_long,
         lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  dplyr::select(name, long, lat)
```

# Bear data
```{r}
bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```

# load and bind data sets
```{r}
tail(both_species)
```

# get terrain map
```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat),
                     zoom = 4, 
                     maptype = "terrain-background")



ggmap(base)
```

# plot first bigfoot layer on map 
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```

# plot bear layer 
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))
```

# plot bigfoot compared to black bear
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```
There seems to be a higher concentration of Gigantopithecus slightly above the state of Nevada. Very similar to the previous graph, just with some minor concentration differences. 

## Part 4: Bathymetry

# load libraries 
```{r}
library(tidyverse)
library(marmap)
```

#convert quake depth to feet
```{r}
min_long <- -170
max_long <- 164
min_lat  <- -42
max_lat  <- -8

earthquakes <- quakes %>% 
   mutate(depth = depth * 3280.84)
```


# Will auto read the saved file if present.
```{r}
fiji <- getNOAA.bathy(lon1 = min_long,
                      lon2 = max_long,
                      lat1 = min_lat,
                      lat2 = max_lat,
                      antimeridian = TRUE,
                      keep = TRUE)
```

# Could also just use autoplot() without the .bathy extension.
```{r}
autoplot.bathy(fiji, 
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE)
```

# Using autoplot, which "knows" to use autoplot.bathy
```{r}
autoplot(fiji, 
         geom = c("raster","contour"),
         size = 0.1,
         na.rm = TRUE) +
  scale_fill_gradient2(low = "dodgerblue4", 
                       mid = "gainsboro", 
                       high = "darkgreen",
                       name = "Depth (ft)") +
  labs(x = "Longitude",
       y = "Latitude") +
  theme(axis.title.y = element_text(angle = 0, 
                                    vjust = 0.5))
```

# create base_map
```{r}
base_map <- autoplot(fiji, 
                     geom = c("raster","contour"),
                     size = 0.1,
                     na.rm = TRUE) +
  scale_fill_gradient2(low = "dodgerblue4", 
                       mid = "gainsboro", 
                       high = "darkgreen",
                       name = "Depth (ft)") +
  labs(x = "Longitude",
       y = "Latitude") +
  theme(axis.title.y = element_text(angle = 0, 
                                    vjust = 0.5))
```

```{r}
base_map + 
  geom_point(data = earthquakes,
             aes(x = long,
                 y = lat))
```

# adjust size to help with crowding 
```{r}
base_map + 
  geom_point(data = earthquakes,
             aes(x = long,
                 y = lat,
                 size = mag),
             alpha = 0.4)
```

# set min and max point sizes for magnitude
```{r}
base_map + 
  geom_point(data = earthquakes,
             aes(x = long,
                 y = lat,
                 size = mag),
             alpha = 0.4) +
  scale_size_continuous(range = c(0.5,3), name = "Magnitude")
```

## now I try it

#Set new global variables
```{r}
min_long <- -90
max_long <- -58
min_lat  <- 8
max_lat  <- 25

earthquakes <- quakes %>% 
   mutate(depth = depth * 3280.84)
```


# Load blennies.csv into a blennies object. 
```{r}
blennies <- read_csv(here::here("Data", "blennies.csv"))
```


#Use getNOAA.bathy() function,Set antimeridian to FALSE,Set keep = TRUE,Save the results to a carib_sea object.
```{r}
carib_sea <- getNOAA.bathy(lon1 = min_long,
                      lon2 = max_long,
                      lat1 = min_lat,
                      lat2 = max_lat,
                      antimeridian = FALSE,
                      keep = TRUE)
```

#Use autoplot to plot the base bathymetric map
```{r}
base_map_2 <- autoplot(carib_sea, 
         geom = c("raster","contour"),
         size = 0.1,
         na.rm = TRUE) +
  scale_fill_etopo(guide = FALSE) +
  labs(x = "Longitude",
       y = "Latitude") +
  theme(axis.title.y = element_text(angle = 0, 
                                    vjust = 0.5))
base_map_2
```

#Add the blennies
```{r}
base_map_2 + 
  geom_point(data = blennies,
             aes(x = long,
                 y = lat, shape = name),color = "red", size = 2.8)+ 
  labs(shape = "Species")+
  theme(legend.position = "bottom")+
  theme(legend.text = element_text(face = "italic"))
base_map_2
```

I really enjoyed this one. It's neat to be able to import data and see it represented on graphs with terrain. 

### et Vóila