---
title: "HW 14: Maps"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load libraries 
```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

## Part 1: outline maps 

### U.S maps

# Outline of the 48 contiguous states
```{r}
usa <- map_data(map = "state")
```

## Outline of just Missouri
```{r}
us_missouri <- map_data(map = "state",
                        region = "missouri")
```

# Outline of Alabama, Florida, Mississippi, and North Carolina
```{r}
us_other <- subset(x = usa, 
                   subset = region %in% c("alabama",
                                          "florida",
                                        "mississippi",
                                    "north carolina"))
```

# lay the first layer of maps 
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray")
```


# add missouri to the map
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black")
```
Missouri has been added. 

# add outline of all 48 states
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black")
```
Full map, with Missouri filled and other 4 outlined.

# Add the filled states first so that the black outlines of all
# states are overlaid on top.
```{r}
us_map <- ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black") + 
  theme_map() +
  coord_fixed(1.3)

us_map
```


# build Missouri map and highlight Scott County
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <- ggplot() + 
  geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") +
  theme_map() +
  coord_fixed(1.3)

mo_map
```

# put Missouri map and us map side by side 
```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5,1))
```
checks out, two maps side by side 

### outline maps: my turn
# highlight missouri on the us map
```{r}
ggplot() + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black")
```
































