---
title: "HW 14: Maps"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load libraries 
```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

## Part 1: outline maps 

### U.S maps

# Outline of the 48 contiguous states
```{r}
usa <- map_data(map = "state")
```

## Outline of just Missouri
```{r}
us_missouri <- map_data(map = "state",
                        region = "missouri")
```

# Outline of Alabama, Florida, Mississippi, and North Carolina
```{r}
us_other <- subset(x = usa, 
                   subset = region %in% c("alabama",
                                          "florida",
                                        "mississippi",
                                    "north carolina"))
```

# lay the first layer of maps 
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray")
```


# add missouri to the map
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black")
```
Missouri has been added. 

# add outline of all 48 states
```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black")
```
Full map, with Missouri filled and other 4 outlined.

# Add the filled states first so that the black outlines of all
# states are overlaid on top.
```{r}
us_map <- ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black") + 
  theme_map() +
  coord_fixed(1.3)

us_map
```


# build Missouri map and highlight Scott County
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <- ggplot() + 
  geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") +
  theme_map() +
  coord_fixed(1.3)

mo_map
```

# put Missouri map and us map side by side 
```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5,1))
```
checks out, two maps side by side 

### outline maps: my turn
# highlight missouri on the us map
```{r}
us_map2 <- ggplot() + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black")
us_map2
```

# highlight two counties in gray and birth county in black
```{r}
missouri <- map_data("county", "missouri")
mo_franklin <- subset(missouri, subregion %in% c("franklin"))
mo_holt <- subset(missouri, subregion %in% c("holt"))
mo_lewis <- subset(missouri, subregion %in% c("lewis"))
mo_map2 <- ggplot() + 
  geom_polygon(data = missouri, aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_franklin, aes(x = long, y = lat, group = group), fill = "black") + 
  geom_polygon(data = mo_holt, aes(x = long, y = lat, group = group), fill = "gray55") + 
  geom_polygon(data = mo_lewis, aes(x = long, y = lat, group = group), fill = "gray55") + 
  coord_fixed(1.3) + 
  theme_map()

mo_map2
```

#combine two maps side by side
```{r}
us_map2 + mo_map2 +plot_layout(ncol = 2, widths = c(1.5,1))
```
Bam! two maps of my birth state and county in black, as well as two additional counties in gray. 


## Part 2: Choropleth maps 

# load library
```{r}
library(sf)
library(tidyverse)
```


# import skin data 
```{r}
skin_cancer_raw <- 
  read_delim(file = here::here("data","skin_cancer.txt"), 
             delim = " ", 
             skip = 7, 
             trim_ws = TRUE)
```

# wrangle data 
```{r}
skin_cancer <- skin_cancer_raw %>% 
  filter(State != "Wash,D.C.") %>% 
  mutate(State = case_when(
    State == "NorthCarolina" ~ "North Carolina",
    State == "SouthCarolina" ~ "South Carolina",
    State == "NorthDakota" ~ "North Dakota",
    State == "SouthDakota" ~ "South Dakota",
    State == "NewHampshire" ~ "New Hampshire",
    State == "NewJersey" ~ "New Jersey",
    State == "NewMexico" ~ "New Mexico",
    State == "MewYork" ~ "New York", # Data has MewYork typo 
    State == "RhodeIsland" ~ "Rhode Island",
    State == "WestVirginia" ~ "West Virginia",
    TRUE ~ State
  ))
```


# calculate average mortality for all 48 states. Then use mutate() to add a relative_mort column, substracting the Mort column from the mean_mort average
```{r}
mean_mort <- mean(skin_cancer$Mort)

skin_cancer_adj <- skin_cancer %>% 
  mutate(relative_mort = Mort - mean_mort)
```


# Used to trim Alaska, Hawaii, and territories from the shapefile.
```{r}
lower_48 <- c("Alabama","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
```


# filter and only keep lower_48 states
```{r}
states <- st_read(here::here("data/cb_2017_us_state_500k.shp"),
                  stringsAsFactors = TRUE)
```
```{r}
states_df <- states %>%
  dplyr::filter(NAME %in% lower_48)
```


# create column called STATEFP
```{r}
str(states_df$STATEFP)
```

# Capture the FIPS levels from states_df
# Use to factor the FIPS and state names in the next step.
```{r}
state_levels <- levels(states_df$STATEFP)
name_levels <- levels(states_df$NAME)
```

```{r}
skin_cancer_adj <- left_join(skin_cancer_adj, fips)

states_df <- left_join(states_df, skin_cancer_adj)
```


#change the STATEFP column to an unordered factor.
```{r}
library(tidyverse)
fips <- read_csv(here::here("data","state_fips.csv")) %>% 
  select(-abbr) %>% 
  rename(STATEFP = FIPS) %>% 
  mutate(STATEFP = factor(STATEFP, levels = state_levels))
```

```{r}
ggplot(states_df) +
  geom_sf(aes(fill = relative_mort)) +
  scale_fill_viridis_c(name = "Skin Cancer,\nRelative Mortality",
                       option = "inferno") +
  coord_sf(crs = st_crs(5070)) +
  theme_bw() +
  theme(legend.position = "bottom")
```
It seems like skin cancer mortality is higher in the southern states, which is expected as they receive more intense sunlight closer to the equator .

## Choropleth map: your turn

# import data
```{r}
lyme_disease_raw<- read_csv(here::here("data", "lyme_disease.csv")) 
```


#### Come back to this ###




## Part 3: Dot distribution maps


# load library
```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(ggmap)
```



# Global vars -------------------------------------------------------------

# Restrict the map to the western US.
```{r}
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(9618973)
```



# Bigfoot data ------------------------------------------------------------
```{r}
bigfoot <- read_csv(here::here("data", "bigfoot.csv"))

bigfoot <- bigfoot %>% 
  filter(long >= min_long & long <= max_long,
         lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  dplyr::select(name, long, lat)
```


# Bear data --------------------------------------------------------------
```{r}

bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```

```{r}
tail(both_species)
```


# get a terrain map of data 
```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat),
                     zoom = 4, 
                     maptype = "terrain-background")



ggmap(base)
```


# add plot point layer for big foot 
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```


# add point layer for black bear
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))
```


# make a final plot for both species 
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```


## make my own distribution map

# change setseed to last 7 digits of SO #

####### Come back to this ############



## Part 4: Bathymetry


# load libraries 
```{r}
library(tidyverse)
library(marmap)
```


#convert quake depth to feet
```{r}
min_long <- -170
max_long <- 164
min_lat  <- -42
max_lat  <- -8

earthquakes <- quakes %>% 
   mutate(depth = depth * 3280.84)
```


# Will auto read the saved file if present.
```{r}
fiji <- getNOAA.bathy (lon1 = min_long,
                      lon2 = max_long,
                      lat1 = min_lat,
                      lat2 = max_lat,
                      antimeridian = TRUE,
                      keep = TRUE)
```

```{r}
autoplot.bathy(fiji, 
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE)
```

















