---
title: "HW 13: Strings and things"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries 
```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```
## Part 1: DNA strings

### import data
```{r}
beta_gliadin_raw <- read_lines(here("Data", "beta_gliadin.txt"),skip=1)
beta_gliadin_raw
```
Data matches the format from the notes

### Replace whitespace and digits with nothing. Printe results 
```{r}
beta_gliadin_cleaned <- str_replace_all(beta_gliadin_raw, "\\d", "") %>%
  str_replace_all("\\s", "")
beta_gliadin_cleaned
```
Matches data in notes and is 56 lines long


### Use 'str_c' to collapse the strings
```{r}
beta_gliadan_collapsed <- str_c(beta_gliadin_cleaned,collapse ="")
beta_gliadan_collapsed
```
Matches line in the notes


### Convert it to upper case with str_to_upper().
```{r}
beta_gliadan_upper <- str_to_upper(beta_gliadan_collapsed)
beta_gliadan_upper
```


### Use str_length() returns the length of a string.
```{r}
str_length(beta_gliadan_upper)
```
The DNA sequence is 3310 characters long 


## Find promoter region 

```{r}
str_count(beta_gliadan_upper, "A.C.T.G")
str_count(beta_gliadan_upper, "ATG")
str_count(beta_gliadan_upper, "TGA|TAA")
```
There is 13 A.C.T.G, 66 "ATG", and 99 "TGA|TAA" codons.



### Assign the two motif sequences to variables caat_box and tata_box  
```{r}
caat_box<- "CCAAT"
tata_box<- "CTATAA"
```


### count the number
```{r}
str_count(beta_gliadan_upper,c(caat_box,tata_box))
```
There are 4 "CCAAT" and 1 "CTATAA"



### Make a search string called caat_seq and count number of sequences 
```{r}
caat_seq<- "CCAAT.{100,150}ATG"
str_count(beta_gliadan_upper, caat_seq)
```
There are 2 possible sequences for a string that begins with "CCAAT" and ends with "ATG"


### Make a searcg string called tata_seq and count the number of sequences 
```{r}
tata_seq<- "CTATAA.{50,100}ATG"
str_count(beta_gliadan_upper, tata_seq)
```
There is 1 sequence.


### Use str_extract_all to extract all of the caat_seq sequences 
```{r}
caat_seq_all <- str_extract_all(beta_gliadan_upper,caat_seq,simplify = TRUE)
str_length(caat_seq_all)
```
The two lengths are 139 and 148.


### Use str_which() to find which caat_seq has the tata_box sequence
```{r}
str_which(caat_seq_all,tata_seq)
```
Promotor  sequence is 1. 

### Print promotor and determine it’s length.
```{r}
promotor<- str_sub(caat_seq_all)
promotor[1]
str_length(promotor)
```
The promotor length is 139 and 148.


## Part 2: Ants

### import ants data
```{r}
ants_data <- read_csv(here("Data", "ants.csv"))
ants_data
```


### Wrangle the data 
```{r}
(ants_data_wrangled <- ants_data %>%
   pivot_longer(names_to = "species",
                values_to = "abundance",
                matches("[A-Z].+\\..+",ignore.case = FALSE)) %>%
   select("year", "species", "abundance") %>%
  mutate(species = str_replace(species,
                                "(.{4}).*\\.(.{4}).*", "\\1\\2"),
          year = factor(year,
                        ordered = TRUE,
                        levels = c(2010, 2011))) %>%
   group_by(year, species) %>%
   summarise(total_abundance = sum(abundance))
   )
```



### Plot the data
```{r}
ggplot(ants_data_wrangled,
       aes(x = reorder(species,total_abundance),
       y = total_abundance),
       group = species) +
  geom_line(color="Gray70")+
  geom_point(aes(shape = year),
             size = 2) +
  coord_flip() +
  labs(x = NULL, y = "Total Abundance", shape = "Year") +
theme_minimal()  
```
This matches the graph from the notes. I forgot to add color=grey70. This caused the dots on the plot to not be connected by the horizontal lines.


## Part 3:  Featuring Phlox Flowers

### Import and wrangle first set of data
```{r}
flower_size_data <- read_csv(here("data","flower_size.csv"))
flower_size_data
```
### wrangle flower data
```{r}
flower_data_wrangled <- flower_size_data %>% 
  filter(!is.na('Flower number')) %>% 
  select(species = 'Species', cor_length = `Corolla length (cm)`, throat_width = `Corolla width throat (cm)`) %>% 
  separate(species, into = c('genus', 'species'), extra = 'drop') %>% 
  group_by(genus, species) %>% 
  summarise(mean_length = mean(cor_length, na.rm = TRUE),
            mean_width = mean(throat_width, na.rm = TRUE),
            max_length = max(cor_length),
            max_width = max(throat_width)
            ,.groups = 'drop')
flower_data_wrangled
```
Wrangled data checks out with the notes. Used drop_na to remove NA flower numbers. 


### import second data set 
```{r}
pollinators_data <- read_csv(here("Data", "flower_pollinators.csv"))
pollinators_data
```


### Remove columns, rows, and seperate species
```{r}
pollinators <- pollinators_data %>% 
  select(-(Color | starts_with("Source"))) %>% 
  drop_na("Pollinator") %>% 
  separate(Species, into = c("genus", "species"), extra = 'drop') %>% 
  separate(Pollinator, into = "Syndrome", extra = 'drop')
pollinators
```


### Join 2 datasets using inner_join to join smaller to larger
```{r}
flower_joined_data <- flower_data_wrangled %>%
  inner_join(pollinators)
flower_joined_data
```
The two data sets are joined 


### Make a scatterplot to show the relationship between mean_length (x-axis) and mean_width (y-axis)
```{r}
flower_joined_data %>% 
  ggplot() +
  geom_point(mapping = aes(x = mean_length, y = mean_width, color = Syndrome)) +
  labs(x = "Mean length", y = "Mean width")
```
Graph seems to match the notes.bats seem to have the largest width and length mean.

### Use mutate with str_replace() and regex patterns to replace all occurences of “beetle” with “beetles,” and “bee” with “bees”.
```{r}
flower_mutated <- flower_joined_data %>%
  mutate(Syndrome = str_replace(Syndrome, "(^be.*e$)", "\\1s")) %>%
  ggplot() +
  geom_point(mapping = aes(x = mean_length, y = mean_width, color = Syndrome)) +
  labs(x = "Mean length", y = "Mean width")
flower_mutated
```
'Beetle' was changed to 'beetles' and 'bee' was changed to 'bees'.


### use mutate() with str_replace and a regex pattern to replace all instances of “beetles” or “flies” with “other insects”.
```{r}
flower_mutated_2 <- flower_joined_data %>% 
 mutate(Syndrome = str_replace(Syndrome, "(^be.*e$)", "\\1s")) %>%
  mutate(Syndrome = str_replace(Syndrome, '(beetles|flies)', 'other insects')) %>% 
  ggplot() +
  geom_point(mapping = aes(x = mean_length, y = mean_width, color = Syndrome)) +
  labs(x = "Mean length", y = "Mean width")
flower_mutated_2
```
This has added another data pointed names "other insects", as well as removing "flies".



### Use mutate and case_when to create a new column called syndrome_group and to change autogamous to Autogamous (upper case A), bat to Mammal, hummingbird to Bird, and everything else to Insect. Creating a new column would allow you to access the individual syndromes for future plots, when necessary.
```{r}
flower_mutated_3 <- flower_joined_data %>%
   mutate(Syndrome = str_replace(Syndrome, "(^be.*e$)", "\\1s")) %>%
  mutate(Syndrome = str_replace(Syndrome, '(beetles|flies)', 'other insects')) %>% 
  mutate(syndrome_group = case_when(
           Syndrome == "autogamous" ~ "Autogamous",
           Syndrome == "bat" ~ "Mammal",
           Syndrome == "hummingbird" ~ "Bird",
           TRUE ~ "Insect"))
flower_mutated_3 %>% 
   ggplot(aes(x=mean_length,y=mean_width,color=syndrome_group))+
  geom_point()+
  labs(x="Mean length",y="Mean width",color="Syndrome")
flower_mutated_3
```
This has created 4 syndromes now and allows fr the data pattern to be seen easier. Mammals have the greatest mean length and wisth, which relfects the bat listed earlier. 

### Highlighting groups with gghighlight
```{r}
flower_mutated_3 %>%
  ggplot(aes(x=mean_length,y=mean_width,color=syndrome_group))+
  geom_point()+
  labs(x="Mean length",y="Mean width",color="Syndrome")+
  gghighlight(syndrome_group=="Mammal"|syndrome_group=="Bird",use_direct_label = FALSE)
```
This highlights the mammals and birds on the graph.


###





































