---
title: "HW 13: Strings and things"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries 
```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```
## Part 1: DNA strings

### import data
```{r}
beta_gliadin_raw <- read_lines(here("Data", "beta_gliadin.txt"),skip=1)
beta_gliadin_raw
```
Data matches the format from the notes

### Replace whitespace and digits with nothing. Printe results 
```{r}
beta_gliadin_cleaned <- str_replace_all(beta_gliadin_raw, "\\d", "") %>%
  str_replace_all("\\s", "")
beta_gliadin_cleaned
```
Matches data in notes and is 56 lines long


### Use 'str_c' to collapse the strings
```{r}
beta_gliadan_collapsed <- str_c(beta_gliadin_cleaned,collapse ="")
beta_gliadan_collapsed
```
Matches line in the notes


### Convert it to upper case with str_to_upper().
```{r}
beta_gliadan_upper <- str_to_upper(beta_gliadan_collapsed)
beta_gliadan_upper
```


### Use str_length() returns the length of a string.
```{r}
str_length(beta_gliadan_upper)
```
The DNA sequence is 3310 characters long 


## Find promoter region 

```{r}
str_count(beta_gliadan_upper, "A.C.T.G")
str_count(beta_gliadan_upper, "ATG")
str_count(beta_gliadan_upper, "TGA|TAA")
```
There is 13 A.C.T.G, 66 "ATG", and 99 "TGA|TAA" codons.



### Assign the two motif sequences to variables caat_box and tata_box  
```{r}
caat_box<- "CCAAT"
tata_box<- "CTATAA"
```


### count the number
```{r}
str_count(beta_gliadan_upper,c(caat_box,tata_box))
```
There are 4 "CCAAT" and 1 "CTATAA"



### Make a search string called caat_seq and count number of sequences 
```{r}
caat_seq<- "CCAAT.{100,150}ATG"
str_count(beta_gliadan_upper, caat_seq)
```
There are 2 possible sequences for a string that begins with "CCAAT" and ends with "ATG"


### Make a searcg string called tata_seq and count the number of sequences 
```{r}
tata_seq<- "CTATAA.{50,100}ATG"
str_count(beta_gliadan_upper, tata_seq)
```
There is 1 sequence.


### Use str_extract_all to extract all of the caat_seq sequences 
```{r}
caat_seq_all <- str_extract_all(beta_gliadan_upper,caat_seq,simplify = TRUE)
str_length(caat_seq_all)
```
The two lengths are 139 and 148.


### Use str_which() to find which caat_seq has the tata_box sequence
```{r}
str_which(caat_seq_all,tata_seq)
```
Promotor  sequence is 1. 

### Print promotor and determine it’s length.
```{r}
promotor<- str_sub(caat_seq_all)
promotor[1]
str_length(promotor)
```
The promotor length is 139 and 148.


## Part 2: Ants

### import ants data
```{r}
ants_data <- read_csv(here("Data", "ants.csv"))
ants_data
```


### Wrangle the data 
```{r}
(ants_data_wrangled <- ants_data %>%
   pivot_longer(names_to = "species",
                values_to = "abundance",
                matches("[A-Z].+\\..+",ignore.case = FALSE)) %>%
   select("year", "species", "abundance") %>%
  mutate(species = str_replace(species,
                                "(.{4}).*\\.(.{4}).*", "\\1\\2"),
          year = factor(year,
                        ordered = TRUE,
                        levels = c(2010, 2011))) %>%
   group_by(year, species) %>%
   summarise(total_abundance = sum(abundance))
   )
```



### Plot the data
```{r}
ggplot(ants_data_wrangled,
       aes(x = reorder(species,total_abundance),
       y = total_abundance),
       group = species) +
  geom_line(color="Gray70")+
  geom_point(aes(shape = year),
             size = 2) +
  coord_flip() +
  labs(x = NULL, y = "Total Abundance", shape = "Year") +
theme_minimal()  
```
This matches the graph from the notes. I forgot to add color=grey70. This caused the dots on the plot to not be connected by the horizontal lines.


## Part 3:  Featuring Phlox Flowers

### Import and wrangle first set of data
```{r}
flower_size_data <- read_csv(here("data","flower_size.csv"))
flower_size_data
```
### wrangle flower data
```{r}
flower_size_wrangled <- flower_size_data %>%
  drop_na("Flower number") %>% 
  select("species"=Species,"cor_length"=`Corolla length (cm)`,"throat_width"=`Corolla width throat (cm)`) %>% 
  separate(species, into = c("genus", "species"), extra = "drop") %>%
   group_by('genus', 'species') %>%
   summarise(mean_length=mean(cor_length,na.rm = TRUE),
            mean_width=mean(throat_width,na.rm = TRUE),
            max_length=max(cor_length,na.rm = TRUE),
            max_width=max(throat_width,na.rm = TRUE),.groups='drop') 
flower_size_data
```
Wrangled data checks out with the notes. Used drop_na to remove NA flower numbers. 


### import second data set 















































