---
title: "HW 13: Strings and things"
author: "Hayley Laubinger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Load Libraries 
```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```
## Part 1: DNA strings

### import data
```{r}
beta_gliadin_raw <- read_lines(here("Data", "beta_gliadin.txt"),skip=1)
beta_gliadin_raw
```
Data matches the format from the notes

### Replace whitespace and digits with nothing. Printe results 
```{r}
beta_gliadin_cleaned <- str_replace_all(beta_gliadin_raw, "\\d", "") %>%
  str_replace_all("\\s", "")
beta_gliadin_cleaned
```
Matches data in notes and is 56 lines long


### Use 'str_c' to collapse the strings
```{r}
beta_gliadan_collapsed <- str_c(beta_gliadin_cleaned,collapse ="")
beta_gliadan_collapsed
```
Matches line in the notes


### Convert it to upper case with str_to_upper().
```{r}
beta_gliadan_upper <- str_to_upper(beta_gliadan_collapsed)
beta_gliadan_upper
```


### Use str_length() returns the length of a string.
```{r}
str_length(beta_gliadan_upper)
```
The DNA sequence is 3310 characters long 


## Find promoter region 

### use str_count() and regex to count the number of IUPAC letters in the sequence that are not A, C, G, or T. 
```{r}
str_count(beta_gliadan_upper, "[A.C.T.G]")
str_count(beta_gliadan_upper, "ATG")
str_count(beta_gliadan_upper, "TGA|TAA")
```









































